---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzip the data and load it into a data frame.
```{r echo=TRUE}
unzip("activity.zip")
activity <- read.csv("activity.csv")
str(activity)

```



## What is mean total number of steps taken per day?
Aggregate the data frame by day, then take the mean.
```{r echo=TRUE}
steps.per.day <- aggregate(activity$steps, 
                           list(day=activity$date), sum)
hist(steps.per.day$x, 
     breaks=10, 
     main="Histogram of steps per day",
     xlab="Steps")
```
```{r echo=TRUE}
```
###Mean
```{r echo=TRUE}
mean(steps.per.day$x, na.rm=TRUE)
```
###Median
```{r echo=TRUE}
median(steps.per.day$x, na.rm=TRUE)

```

## What is the average daily activity pattern?

```{r}
avg.per.interval <- aggregate(activity$steps, 
                           list(Interval=activity$interval), 
                           function(steps){
                               mean(steps, na.rm=TRUE)
                           })

plot(avg.per.interval$Interval,
     avg.per.interval$x,
     type='l',
     main="Average daily activity",
     xlab="Interval (minutes)",
     ylab="Avg number of steps")


```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avg.per.interval[which.max(avg.per.interval$x),"Interval"]
```

## Imputing missing values

+ Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(!complete.cases(activity))

```
+ Devise a strategy for filling in all of the missing values in the dataset. (We will use the mean for the 5 minute interval)

```{r}
activity$steps <- lapply(seq_along(activity$steps), function(i) {
    if(is.na(activity$steps[[i]])) {
        avg.per.interval$x[[which(avg.per.interval$Interval == activity$interval[[i]])]]
    } else {
        activity$steps[[i]]
    }
})
activity[140:145,]

```


## Are there differences in activity patterns between weekdays and weekends?
